---
title: 'An Exploratory Analysis of Corruption and parking violations'
author: "Shiraz Chakraverty"
date: "May 15, 2018"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
\begin{center}
Exploratory data analysis of the relationship between corruption and parking violations both pre and post 2002 and any 
other relevant explanatory variables. The dataset is relatively small at 365 records. There are 28 variables. We have a set of metric 
and categorical variables. The categorical variables predominantly identify the country...

Data Issues: 
1) na values in a range of 64-68 rows: We are unable to use partial elements, as that will give us invalid computations. Thus these rows are being removed.
2) Category variables : 
3) 
4) 
\end{center}

## Data Selection Strategy
--Add comments here to describe data selection strategy. Explain upper and lower bounds of the dataset, explain choice of variables.



## Setup

First, we load the car library, which gives us a convenient list of function.

```{r car}
library(car)
```


```{r loading}
# Load the data 
getwd()
load("Corrupt.Rdata")
```

## Data Selection



We note that we have `r nrow(Fert_all)` observations and `r length(colnames(Fert_all))` variables.

```{r shape}
nrow(FMcorrupt)
str(FMcorrupt)
```

The base dataset has about 68 rows with na data elements.

```{r remove_na}
corrupt_v1 = FMcorrupt[complete.cases(FMcorrupt),]
summary(corrupt_v1)
```

##Univariate Analysis
Use visualizations and descriptive statistics to perform a univariate analysis of each key variable. Be sure to
describe any anomalies, coding issues, or potentially erroneous values. Explain how you respond to each issue
you identify. Note any features that appear relevant to statistical analysis. Discuss what transformations
may be appropriate for each variable.


```{r}
```

### Analysis and feature review -- tbd
Our first step is to study each variable closely, understand the distribution, measures of location and create a few graphs
to understand the dataset better. Here we are also going to look for issues in the values that may require further cleanup. As 
we make findings, we will update this section with our findings and note coding repairs we make to fix.
the data.  One primary seperation we will do here is the look at the pre and post 2002 voilations and records seperately, this 
gives us another perspective into the effects of removal of diplomatic immunity in the area of parking citations.


```{r}
corrupt_v1_pre = subset(corrupt_v1, prepost == "pre")
nrow(corrupt_v1_pre)
corrupt_v1_post = subset(corrupt_v1, prepost == "pos")
nrow(corrupt_v1_post)
```

Create graphs to look at distributions of metric and category variables

```{r}

```

Our post 2002 data frame has just `r nrow(corrupt_v1_post)` observations.
Our pre 2002 data frame has just `r nrow(corrupt_v1_pre)` observations.
We examine the respective summary.

```{r}
summary(corrupt_v1_pre)
summary(corrupt_v1_post)
```

## Exploratory Analysis -tbd

Analysis of Key Relationships
Explore how your outcome variable is related to the other variables in your dataset. Make sure to use
visualizations to understand the nature of each bivariate relationship.
What tranformations can you apply to clarify the relationships you see in the data? Be sure to justify each
transformation you use.

```{r}
```


Analysis of Secondary Effects (10 pts) -- tbd
What secondary variables might have confounding effects on the relationships you have identified? Explain
how these variables affect your understanding of the data.

Conclusion (20 pts) -- tbd
Summarize your exploratory analysis. What can you conclude based on your analysis?

